---
title: "Extreme heat effects on monarchs and OE"
author: "Isabella Ragonese"
date: "12/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load libraries, include=FALSE}
library(tidyverse) 
library(ggplot2) 
library(lme4)
library(visreg)
library(lme4)
library(plyr)
library(mgcv)
library(tidymv)
library(voxel)
library(MASS)
library(mgcViz)
library(itsadug)
library(sjPlot)
```

#Bolker 
https://bbolker.github.io/mixedmodels-misc/glmmFAQ#overdispersion 
#overdispersion function
```{r}
overdisp_fun <- function(model) {
    rdf <- df.residual(model)
    rp <- residuals(model,type="pearson")
    Pearson.chisq <- sum(rp^2)
    prat <- Pearson.chisq/rdf
    pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
    c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}
```

#other sources: 
https://noamross.github.io/gams-in-r-course/
https://jacolienvanrij.com/Tutorials/GAMM.html

#read in data and create subsets
```{r, read in data and create subsets for control monarchs, pupal and larval immune measures, include=FALSE}
mtemp.comp<-read.csv("data/mtemp_comp_Nov1_2021.csv") #master data

mtemp.comp$Spore_load2<-ifelse(mtemp.comp$Infection_status != 0, mtemp.comp$Spore_load, 0) #add column that includes 0 spore load for adults that emerged without parasites
mtemp.comp$inoc_status<-ifelse(mtemp.comp$OE_strain != "control", 1, 0) #add column that represents whether monarch was inoculated with either strain of OE

#create factor variables for GAM
mtemp.comp$Lin_f<-as.factor(mtemp.comp$Lineage)
mtemp.comp$Sex_f<-as.factor(mtemp.comp$Sex)
mtemp.comp$inoc_f<-as.factor(mtemp.comp$inoc_status)
mtemp.comp$OE_f<-as.factor(mtemp.comp$OE_strain)

mtemp.comp$Wing_deformity_bin<-if_else(mtemp.comp$Wing_deformity==0|mtemp.comp$Wing_deformity==1,0,1) #make binary deformed/not deformed instead on the 0-4 scale

#comprehensive dataframe with just the infected individuals
mtemp.inf <- subset(mtemp.comp, Infection_status==1)
#comprehensive dataframe with just the inoculated individuals
mtemp.inoc <- subset(mtemp.comp, inoc_status==1)
#comprehensive dataframe with just control individuals
mtemp.control <- subset(mtemp.comp, OE_strain=="control", Infection_status!=1) 

#subset comprehensive dataframe to create dataframes with either larval or pupal bled individuals
pupahemo.df = filter(mtemp.comp, Bleed_stage.x=="Pupa")
#Add log10 Hemo for ID#70 - as if 1 cell found
pupahemo.df[28,30]=1.130334

larvahemo.df = filter(mtemp.comp, Bleed_stage.x=="Larva")

#take out rows with NA for loghemo, or final absorbance (some individuals didn't have enough hemolymph for both analyses) in larvahemo, pupalhemo, and po datasets
completeFun <- function(data, desiredCols) {
  completeVec <- complete.cases(data[, desiredCols])
  return(data[completeVec, ])
}

larvahemo.df<-completeFun(larvahemo.df, "LogHemo")
larvahemo.df = filter(larvahemo.df, !is.na(Mass))
pupahemo.df<-completeFun(pupahemo.df, "LogHemo")
pupahemo.df = filter(pupahemo.df, !is.na(Mass))

mtemp.po<-completeFun(mtemp.comp, "FinalAbsorbance")
mtemp.adult<-filter(mtemp.comp, Surv_adult==1)
control.adult<-completeFun(mtemp.control, "Dev_adult")

#merozoites data #n=46
merozoite_df<-read.csv("data/Merozoites_7_21.csv")
merozoite_df2<-left_join(merozoite_df, mtemp.inf, by="ID")
merozoite_df2$Temp<-as.character(merozoite_df2$Temp)
merozoite_df3<-merozoite_df2[!is.na(merozoite_df2$Temp),]
```

############################
#preliminary visualizations and analyses to asses how bleed treatment affects response variables
## Which response variables were impacted by bleed treatment?
```{r, bleeding effects, echo=FALSE}
Bleed_stage <- mtemp.comp$Bleed_stage.x
response<-mtemp.comp$Wing_deformity_bin
Mod_Bleed<-lm(response ~ Bleed_stage, data = mtemp.comp)
summary(Mod_Bleed)
visreg(Mod_Bleed, ylab="wing deformity")

response<-mtemp.comp$Dev_pupa
Mod_Bleed<-glm(response ~ Bleed_stage, data = mtemp.comp, family=poisson)
summary(Mod_Bleed)
visreg(Mod_Bleed, ylab="Time to pupation")

response<-mtemp.comp$Dev_adult
Mod_Bleed<-glm(response ~ Bleed_stage, data = mtemp.comp, family=poisson)
summary(Mod_Bleed)
visreg(Mod_Bleed, ylab="time to eclosion")

response<-mtemp.comp$Surv_pupa
Mod_Bleed<-glm(response ~ Bleed_stage, data = mtemp.comp, family=binomial)
summary(Mod_Bleed)
visreg(Mod_Bleed, ylab="survival to pupaption")

response<-mtemp.comp$Surv_adult
Mod_Bleed<-glm(response ~ Bleed_stage, data = mtemp.comp, family = binomial)
summary(Mod_Bleed)
visreg(Mod_Bleed, ylab="survival to eclosion")

response<-mtemp.comp$Area
Mod_Bleed<-lm(response ~ Bleed_stage, data = mtemp.comp)
summary(Mod_Bleed)
visreg(Mod_Bleed, ylab="wing area")

response<-mtemp.comp$Adult_longevity
Mod_Bleed<-glm(response ~ Bleed_stage, data = mtemp.comp, family = poisson)
summary(Mod_Bleed)
visreg(Mod_Bleed, ylab="adult longevity")

response<-mtemp.comp$Infection_status
Mod_Bleed<-glm(response ~ Bleed_stage, data = mtemp.comp, family = binomial)
summary(Mod_Bleed)
visreg(Mod_Bleed, ylab="infection (y/n)")

response<-mtemp.comp$logSpore_load
Mod_Bleed<-lm(response ~ Bleed_stage, data = mtemp.comp)
summary(Mod_Bleed)
visreg(Mod_Bleed, ylab="log spore load")
```


################################
#control monarchs
##Development time
```{r}
response<-log10(control.adult$Dev_adult)

m1_mixed<-gam(response ~ s(Temp, k=5) + Sex_f + s(Lin_f,Temp, bs="re"), data=control.adult, method = "REML") #mixed with lineage as random slope; Using factor versions of predictor variables

gam.check(m1_mixed)

summary(m1_mixed)
anova.gam(m1_mixed)

#basic plot
plot(m1_mixed, residuals = TRUE, pch =1, cex=0.8, shade=TRUE, shift = coef(m1_mixed)[1], seWithMean = TRUE)


#add itsadug points- need to put all lines into the console
plot.new()
par(mai = c(1, 1, 1, 1))
plot_smooth(m1_mixed, view = "Temp", rm.ranef = FALSE, col="darkgoldenrod3", lwd=2, rug = TRUE, ylab=expression("Development time "(log[10](days))), xlab=("Temperature (°C)"), hide.label = TRUE) 
plot_data(m1_mixed, view = "Temp", input="data", rm.ranef = FALSE, add = TRUE, alpha=0.5, col="black", hide.label = TRUE)
axis(1, at=c(15,35), labels=FALSE, lwd.ticks=0)
#axis(1, at=seq(18 , 34, by=4), lwd=0, lwd.ticks=1)


```
##Survival
binary response variable = Survival to adult (could do same for pupa)  
  explanatory variables = Temperature
  random = Lineage
  distribution = binomial
```{r}
control.surv<-completeFun(mtemp.control, "Surv_adult")
response<-control.surv$Surv_adult

m1_mixed<-gam(response ~ s(Temp, k=5) + s(Lin_f, Temp, bs="re"), data=control.surv, method = "REML", family=binomial) #lineage as random slope; s(Lin_f, bs="re") would be random intercept

summary(m1_mixed)
gam.check(m1_mixed)
anova.gam(m1_mixed)

##################################
plot(m1_mixed, shade=TRUE, trans = plogis, shift = coef(m1_mixed)[1], seWithMean = TRUE)

#add itsadug points- need to put all lines into the console
plot.new()
par(mai = c(1, 1, 1, 1))
plot_smooth(m1_mixed, view = "Temp", rm.ranef = FALSE, transform = plogis, col="darkgoldenrod3", lwd=2, rug = TRUE, ylab=("Survival Probability"), xlab=("Temperature (°C)"), hide.label = TRUE)
axis(1, at=c(15,35), labels=FALSE, lwd.ticks=0)

#table for prob and standard error points
mse_survprob<-
  control.surv %>% 
  group_by(Temp) %>% 
  dplyr::summarise(propsurv = sum(as.numeric(Surv_adult), na.rm = T)/n(),
                   sepropsurv = sqrt(propsurv*(1-propsurv)/n()),
                   n=n())

x<-mse_survprob$Temp
y<-mse_survprob$propsurv
se<-mse_survprob$sepropsurv

plot.new()
par(mai = c(1, 1, 1, 1))
plot_smooth(m1_mixed, view = "Temp", rm.ranef = FALSE, transform = plogis, col="darkgoldenrod3", lwd=2, rug = TRUE, ylab=("Survival Probability"), xlab=("Temperature (°C)"), ylim=c(0.4,1), hide.label = TRUE)
axis(1, at=c(15,35), labels=FALSE, lwd.ticks=0)
points(x,y,pch=19)
arrows(x,y-se,x,y+se,length = 0.05, angle = 90,code = 3)

```

##Monarch Size
Continuous response variable = Area (n=92) OR Mass (pupa n=104 and larva n=116)
  explanatory variables = Temperature, (sex and bleed stage for Area)
  random = Lineage
  distribution = gaussian  
```{r}
control.adult2<-filter(control.adult, !is.na(Area))
Bleed_stage_f<-as.factor(control.adult2$Bleed_stage.x)
response<-control.adult2$Area

m1_mixed<-gam(response ~ s(Temp, k=5) + Sex_f + Bleed_stage_f + s(Lin_f, Temp, bs="re"), data=control.adult2, method = "REML") #lineage as random effect

summary(m1_mixed)
gam.check(m1_mixed)
anova.gam(m1_mixed)


plot(m1_mixed, shade=TRUE, shift = coef(m1_mixed)[1], seWithMean = TRUE)

#add itsadug points- need to put all lines into the console
plot.new()
par(mai = c(1, 1, 1, 1))
plot_smooth(m1_mixed, view = "Temp", rm.ranef = FALSE, col="darkgoldenrod3", lwd=2, rug = TRUE, ylab=bquote("Wing Area " (mm^2)), xlab=("Temperature (°C)"), hide.label = TRUE)
plot_data(m1_mixed, view = "Temp", input="data", rm.ranef = FALSE, add = TRUE, alpha=0.5, col="black", hide.label = TRUE)
axis(1, at=c(15,35), labels=FALSE, lwd.ticks=0)

#table of monarchs per temp
tab1<-control.adult2 %>% group_by(Temp) %>% dplyr::summarise(total_wings=n())
#could run with the whole dataset and color points by infection status

```

##Adult longevity
```{r}
#response<-log10(control.adult$Adult_longevity)#log it to use gaussian distribution - gam check suggests poor fit
response<-(control.adult$Adult_longevity) #Neg binomial seems to work best looking at gam check

m1_mixed<-gam(response ~ s(Temp, k=5) + Sex_f + s(Lin_f, Temp, bs="re"), data=control.adult, method = "REML", family = 'nb', link='log') #mixed with lineage as random effect

overdisp_fun(m1_mixed) #poisson was overdispersed, looks ok with neg-bin
summary(m1_mixed)
gam.check(m1_mixed)

#plot(m1_mixed)
plot(m1_mixed, residuals = TRUE, pch =1, cex=0.8, shade=TRUE, shift = coef(m1_mixed)[1], seWithMean = TRUE)

#add itsadug points- need to put all lines into the console
#with negative binomial model - use transform = exp to graph data without the log link
plot.new()
par(mai=c(1,1,1,1))
plot_smooth(m1_mixed, view = "Temp", transform=exp, rm.ranef = FALSE, col="darkgoldenrod3", lwd=2, rug = TRUE, ylab=("Adult longevity (days)"), xlab=("Temperature (°C)"), hide.label = TRUE)
plot_data(m1_mixed, view = "Temp", input="data", rm.ranef = FALSE, add = TRUE, alpha=0.5, col="black", hide.label = TRUE)
axis(1, at=c(15,35), labels=FALSE, lwd.ticks=0)

```

###Infection outcomes

##infection success
response variable (binary) = yes/no infection 
  explanatory variables = Temperature, sex, OE strain? size? 
  random= Lineage
  distribution = binomial 
```{r}
mtemp.inoc1 = mtemp.inoc %>% filter(!is.na(Sex))
response<-mtemp.inoc1$Infection_status

m1_mixed<-gam(response ~ s(Temp, k=5) + Sex_f + OE_f + s(Lin_f, Temp, bs="re"), data=mtemp.inoc1, method = "REML", family=binomial) #lineage as random effect

summary(m1_mixed)
gam.check(m1_mixed)

plot(m1_mixed, shade=TRUE, trans = plogis, shift = coef(m1_mixed)[1], seWithMean = TRUE)

meanerror_infsuccess<-
  mtemp.inoc1 %>% 
  group_by(Temp) %>% 
  dplyr::summarise(propinf = sum(as.numeric(Infection_status),na.rm = T)/n(),
                   sepropinf = sqrt(propinf*(1-propinf)/n()),
                   n=n())
#error bar plot
x<-meanerror_infsuccess$Temp
y<-meanerror_infsuccess$propinf
se<-meanerror_infsuccess$sepropinf

plot.new()
par(mai = c(1, 1, 1, 1))
plot_smooth(m1_mixed, view = "Temp", rm.ranef = FALSE, transform = plogis, col="skyblue2", lwd=2, rug = FALSE, ylab=("Infection Probability"), xlab=("Temperature (°C)"), hide.label = TRUE)
points(x,y,
     ylim=c(0,1),
     pch=19)
arrows(x, y-se, x, y+se, length=0.05, angle=90, code = 3)
axis(1, at=c(15,35), labels=FALSE, lwd.ticks=0)

#add itsadug points- need to put all lines into the console
plot.new()
par(mai = c(1, 1, 1, 1))
plot_smooth(m1_mixed, view = "Temp", rm.ranef = FALSE, transform = plogis, col="skyblue2", lwd=2, rug = FALSE, ylab=("Infection Probability"), xlab=("Temperature (°C)"), hide.label = TRUE)
axis(1, at=c(15,35), labels=FALSE, lwd.ticks=0)
```

##Spore load 18-30
response variable = log10 spore load
  explanatory variables = Temperature, sex, 
                (size? (will exclude deformed ones if I use area))
  random= Lineage
  distribution = gaussian
```{r}
mtemp.inf2<-dplyr::filter(mtemp.inf, !is.na(Sex_f)) #18-30C monarchs - infected only
response<-log10(mtemp.inf2$Spore_load)

m1_mixed<-gam(response ~ s(Temp, k=4) + Sex_f + OE_f + s(Lin_f,Temp, bs="re"), data=mtemp.inf2, method = "REML") #lineage as random effect, (binary)

summary(m1_mixed)
gam.check(m1_mixed)
anova.gam(m1_mixed)

plot(m1_mixed, shade=TRUE, shift = coef(m1_mixed)[1], seWithMean = TRUE)

```


##Merozoite concentration - included in supplemental due to small sample size
All from temps 18-30C
  response variable = log10 merozoite concentration
  explanatory variables = Temperature, OE_strain, Mass
  random= Lineage
  distribution = gaussian (zero inflated? Instead, included only 46 samples that were bled at appropriate time)
```{r}
merozoite_df4<-filter(merozoite_df3, !is.na(Mass))
merozoite_df4$Temp<-as.integer(merozoite_df4$Temp)

response<-merozoite_df4$LogOE

m1_mixed<-gam(response ~ s(Temp, k=4) + OE_f + s(Mass) +
                s(Lin_f, Temp, bs="re"), 
              data=merozoite_df4, method = "REML") #lineage as random effect

summary(m1_mixed) #Slight decrease in merozoite concentration at 30C
gam.check(m1_mixed)

#add itsadug points- need to put all lines into the console
plot.new()
par(mai = c(1, 1, 1, 1))
plot_smooth(m1_mixed, view = "Temp", rm.ranef = FALSE, col="skyblue2", lwd=2, rug = TRUE, ylab=bquote("Merozoite concentration "(log[10](cells/ul))), xlab=("Temperature (°C)"), hide.label = TRUE)
plot_data(m1_mixed, view = "Temp", input="data", rm.ranef = FALSE, add = TRUE, alpha=0.5, col="black", hide.label = TRUE)
axis(1, at=c(15,35), labels=c("",""), lwd.ticks=0)
```

##Wing deformity
# Control
```{r}
control.adult3<-filter(control.adult, !is.na(Wing_deformity))
control.adult3$Bleed_stage_f<-as.factor(control.adult3$Bleed_stage.x)
response<-as.factor(control.adult3$Wing_deformity_bin)


m1_mixed<-gam(response ~ s(Temp, k=5) + Sex_f + s(Lin_f, Temp, bs="re"), data=control.adult3, method = "REML", family=binomial) #lineage as random effect


summary(m1_mixed) 
gam.check(m1_mixed)

deformitytable<-control.adult3 %>% 
  group_by(Temp) %>% 
  dplyr::summarise(prop_deform = sum(as.numeric(Wing_deformity_bin), na.rm=T)/n(),
            seprop_deform = sqrt(prop_deform*(1-prop_deform)/n()),
            n = n())

x<-deformitytable$Temp
y<-deformitytable$prop_deform
se<-deformitytable$seprop_deform

plot.new()
par(mai = c(1, 1, 1, 1))
plot_smooth(m1_mixed, view = "Temp", rm.ranef = FALSE, transform = plogis, col="darkgoldenrod3", lwd=2, rug = TRUE, ylab=("Probability of Wing Deformity"), xlab=("Temperature (°C)"), hide.label = TRUE)
axis(1, at=c(15,35), labels=FALSE, lwd.ticks=0)
points(x,y,pch=19,ylim = c(-0.2,0.6))
arrows(x,y-se,x,y+se,length = 0.05, angle = 90, code = 3)

#add itsadug - need to put all lines into the console
plot.new()
par(mai = c(1, 1, 1, 1))
plot_smooth(m1_mixed, view = "Temp", rm.ranef = FALSE, transform = plogis, col="darkgoldenrod3", lwd=2, rug = TRUE, ylab=("Probability of Wing Deformity"), xlab=("Temperature (°C)"), hide.label = TRUE)
axis(1, at=c(15,35), labels=FALSE, lwd.ticks=0)

```


#Immune measures

##Hemocyte concentration - larvae
response variable (continuous) = Log(Larval Hemocytes)  
  explanatory variables = Temperature, OE(y/n), Temp-OE interaction, size, lineage
  distribution = gaussian
```{r}
response<-larvahemo.df$LogHemo

m1_mixed<-gam(response ~ s(Temp,k=5) + inoc_f + s(Mass) + s(Lin_f, Temp, bs="re"), data=larvahemo.df, method = "REML")

summary(m1_mixed) #Temperature*inoculation interaction not sig, neither is temperature #spline for mass is highly significant - increasing mass = increasing hemocytes up to a point
gam.check(m1_mixed)

plot(m1_mixed, shade=TRUE, shift = coef(m1_mixed)[1], seWithMean = TRUE)

#add itsadug points- need to put all lines into the console
plot.new()
par(mai = c(1, 1, 1, 1))
plot_data(m1_mixed, view = "Temp", input="data", rm.ranef = FALSE, alpha=0.7, col="black", main="", ylab=bquote("Larval hemocycte conc. "(log[10](cells/ul))), xlab=("Temperature (°C)"), hide.label = TRUE)
plot_smooth(m1_mixed, view = "Temp", add = TRUE, rm.ranef = FALSE, col="#009E73", lwd=2, rug = TRUE, hide.label = TRUE)
axis(1, at=c(15,35), labels=c("",""), lwd.ticks=0)

#Mass
plot.new()
par(mai = c(1, 1, 1, 1))
plot_data(m1_mixed, view = "Mass", input="data", rm.ranef = FALSE, alpha=0.5, col="black", ylab=bquote("Larval hemocycte conc. "(log[10](cells/ul))), xlab=("Mass (g)"), hide.label = TRUE)
plot_smooth(m1_mixed, view = "Mass", add = TRUE, rm.ranef = FALSE, col="#009E73", lwd=2, rug = TRUE, hide.label = TRUE)
axis(1, at=c(0,2.0), labels=c("",""), lwd.ticks=0)


```

##Hemocyte conc - pupae
response variable (continuous) = Log(pupal Hemocytes)  
  explanatory variables = Temperature, OE(y/n), Temp-OE interaction, size, lineage
  distribution = gaussian
```{r}
response<-pupahemo.df$LogHemo
#response<-pupahemo.df$Hemo_ul #gam.check better with hemo_ul not transformed?

m1_mixed<-gam(response ~ s(Temp, k=5) + inoc_f + s(Mass) + s(Lin_f, Temp, bs="re"), data=pupahemo.df, method = "REML")

summary(m1_mixed) #Temperature smooth significant - decreasing pupal hemocytes with increasing temperature in all pupae (control and inoculated - interaction not sig). Mass not important at this life stage
gam.check(m1_mixed)
#gam.check(m2_mixed)

plot(m1_mixed, shade=TRUE, shift = coef(m1_mixed)[1], seWithMean = TRUE)

#add itsadug points- need to put all lines into the console
plot.new()
par(mai = c(1, 1, 1, 1))
plot_data(m1_mixed, view = "Temp", input="data", main="", rm.ranef = FALSE, alpha=0.5, col="black", ylab=bquote("Pupal hemocycte conc. "(log[10](cells/ul))), xlab=("Temperature (°C)"), hide.label = TRUE)
plot_smooth(m1_mixed, view = "Temp", add = TRUE, rm.ranef = FALSE, col="#009E73", lwd=2, rug = TRUE, hide.label = TRUE)
axis(1, at=c(15,35), labels=c("",""), lwd.ticks=0)
```

##Larval PO activity 
response variable (continuous) = Final (max) Absorbance of PO assay 
  explanatory variables = Temperature, OE(y/n), Temp-OE interaction, Mass, Lineage  
  distribution = gaussian 
```{r}
response<-larvahemo.df$FinalAbsorbance

m1_mixed<-gam(response ~ s(Temp, k=5) + inoc_f + s(Mass) + s(Lin_f, Temp, bs="re"), data=larvahemo.df, method = "REML")

summary(m1_mixed) #Just the smooth term for mass is significant - increase mass= increased absorbance
gam.check(m1_mixed)


#add itsadug points- need to put all lines into the console
plot.new()
par(mai = c(1, 1, 1, 1))
plot_data(m1_mixed, view = "Temp", input="data", main= "",rm.ranef = FALSE, alpha=0.5, col="black", ylab=("Larval PO activity (absorbance)"), xlab=("Temperature (°C)"), hide.label = TRUE)
plot_smooth(m1_mixed, view = "Temp", add = TRUE, rm.ranef = FALSE, col="#009E73", lwd=2, rug = TRUE, hide.label = TRUE)
axis(1, at=c(15,35), labels=c("",""), lwd.ticks=0)

#mass
plot.new()
plot_data(m1_mixed, view = "Mass", input="data",main="", rm.ranef = FALSE, alpha=0.5, col="black", ylab=("Larval PO activity (absorbance)"), xlab=("Mass (g)"), hide.label = TRUE)
plot_smooth(m1_mixed, view = "Mass", add = TRUE, rm.ranef = FALSE, col="#009E73", lwd=2, rug = TRUE, hide.label = TRUE)
axis(1, at=c(0,2.0), labels=c("",""), lwd.ticks=0)

```

##pupal PO activity 
response variable (continuous) = Final (max) Absorbance of PO assay 
  explanatory variables = Temperature, OE(y/n), Temp-OE interaction, Mass, Lineage  
  distribution = gaussian 
```{r}
response<-pupahemo.df$FinalAbsorbance

m1_mixed<-gam(response ~ inoc_f + s(Temp, k=5) + s(Mass) + s(Lin_f, Temp, bs="re"), data=pupahemo.df, method = "REML")


summary(m1_mixed) #Temp smooth for both inoc statuses is important - general decline in PO with increasing temperature, perhaps slightly elevated again for control pupae at 34C. Inoc status not significant. Lineage smooth term is.
gam.check(m1_mixed)

#add itsadug points- need to put all lines into the console

plot.new()
#plot data not working??
par(mai = c(1, 1, 1, 1))
plot_data(m1_mixed, view = "Temp", input="data", rm.ranef = FALSE, alpha=0.5, col="black", main="", ylab=("Pupal PO activity (absorbance)"), xlab=("Temperature (°C)"), hide.label = TRUE)
plot_smooth(m1_mixed, view = "Temp", add = TRUE, rm.ranef = FALSE, col="#009E73", lwd=2, rug = TRUE, hide.label = TRUE)
axis(1, at=c(15,35), labels=c("",""), lwd.ticks=0)

plot.new()
par(mai = c(1, 1, 1, 1))
plot(pupahemo.df$Temp, pupahemo.df$FinalAbsorbance, cex=0.5, col="gray24", pch=16, ylab=("Pupal PO activity (absorbance)"), xlab=("Temperature (°C)"), bty="n")
plot_smooth(m1_mixed, view = "Temp", add= TRUE, rm.ranef = FALSE, col="#009E73", lwd=2, rug = TRUE, hide.label = TRUE)
axis(1, at=c(15,35), labels=c("",""), lwd.ticks=0)


```


#Costs of infection
#use full dataset in GAMs for longevity, development, survival prob, wing area, prob deformity (done)

# wing deformity with full dataset - monarch fitness with infection as a covariate (cost of infection?)
```{r}
mtemp.adult1<-filter(mtemp.adult, !is.na(Wing_deformity))
mtemp.adult1$Bleed_stage_f<-as.factor(mtemp.adult1$Bleed_stage.x)
mtemp.adult1$infection_f<-as.factor(mtemp.adult1$Infection_status)
response<-as.factor(mtemp.adult1$Wing_deformity_bin)

#inoc status
m1_mixed<-gam(response ~ s(Temp, k=5) + s(Temp, k=5, by=inoc_status) + Sex_f +s(Lin_f, Temp, bs="re"),  data=mtemp.adult1, method = "REML", family=binomial)
 
summary(m1_mixed) #temperature-deformity relationship is not different based on inoculation status
gam.check(m1_mixed)

#spore load
mtemp.adult2<-filter(mtemp.adult1, !is.na(Spore_load)) #look at just the monarchs with values for spore load
response<-as.factor(mtemp.adult2$Wing_deformity_bin)

m2_mixed<-gam(response ~ s(Temp, k=4) + s(logSpore_load) + Sex_f +s(Lin_f, Temp, bs="re"),  data=mtemp.adult2, method = "REML", family=binomial)

summary(m2_mixed) #temperature-deformity relationship is not different based on inoculation status
gam.check(m2_mixed)

##############################

deformitytable<-mtemp.adult1 %>% 
  group_by(Temp) %>% 
  dplyr::summarise(prop_deform = sum(as.numeric(Wing_deformity_bin), na.rm=T)/n(),
            seprop_deform = sqrt(prop_deform*(1-prop_deform)/n()),
            n = n())

#ggplot
ggplot(data=mtemp.adult2, aes(x = logSpore_load, y = Wing_deformity_bin)) +
  labs(x = "spore load", y = "Probability of Wing Deformity") +
  stat_smooth(method = 'gam', method.args = list(family = "binomial"), formula = y ~ s(x, k=4)) +
  theme_classic()

#add itsadug - need to put all lines into the console
plot.new()
plot_smooth(m1_mixed, view = "Temp", plot_all = "inoc_status", rm.ranef = FALSE, transform = plogis, col=c("purple","black"), lwd=2, rug = TRUE, ylab=("Probability of Wing Deformity"), xlab=("Temperature (°C)"), hide.label = TRUE)
axis(1, at=c(15,35), labels=FALSE, lwd.ticks=0)

#add itsadug - need to put all lines into the console
plot.new()
plot_smooth(m2_mixed, view = "logSpore_load", rm.ranef = FALSE, transform = plogis, col="skyblue2", lwd=2, rug = TRUE, ylab=("Probability of Wing Deformity"), xlab=bquote(Log[10] ("Spore load of infected monarchs")), hide.label = TRUE)
axis(1, at=c(15,35), labels=FALSE, lwd.ticks=0)

```

##Development time
```{r}
response<-log10(mtemp.adult$Dev_adult)

#inoculation status n=332
m1_mixed<-gam(response ~ s(Temp, k=5) + s(Temp, k=5, by=inoc_status) + Sex_f + s(Lin_f,Temp, bs="re"), data=mtemp.adult, method = "REML") #Temp, lineage, sex = sig; inoc_status - no sig interaction with temp
gam.check(m1_mixed)
summary(m1_mixed)

#spore load n=175
mtemp.adult2<-filter(mtemp.adult, !is.na(Spore_load))
mtemp.adult2$intensity4<-log10(mtemp.adult2$Spore_load+1)#intensity for the 4 lower temps (others NA)

response<-log10(mtemp.adult2$Dev_adult)
m2_mixed<-gam(response ~ s(Temp, k=4) + s(intensity4) + Sex_f + s(Lin_f,Temp, bs="re"), data=mtemp.adult2, method = "REML") #no effect of spore load

gam.check(m2_mixed)
summary(m2_mixed)

#add itsadug points- need to put all lines into the console
plot.new()
par(mai=c(1,1,1,1))
plot_smooth(m1_mixed, view = "Temp", plot_all = "inoc_status", rm.ranef = FALSE, col=c("purple", "black"), lwd=2, rug = TRUE, ylab=bquote("Development time "(log[10](days))), xlab=("Temperature (°C)"), hide.label = TRUE)
axis(1, at=c(15,35), labels=FALSE, lwd.ticks=0)
#axis(1, at=seq(18 , 34, by=4), lwd=0, lwd.ticks=1)

plot.new()
par(mai=c(1,1,1,1))
plot_smooth(m2_mixed, view = "intensity4",  rm.ranef = FALSE, lwd=2, rug = TRUE, ylab=bquote("Development time "(log[10](days))), xlab=("log spore load"), hide.label = TRUE)
axis(1, at=c(15,35), labels=FALSE, lwd.ticks=0)
#axis(1, at=seq(18 , 34, by=4), lwd=0, lwd.ticks=1)


###################
#dev to pupa
# mtemp.adult2<-filter(mtemp.adult, !is.na(Spore_load2))
# mtemp.adult2$intensity<-log10(mtemp.adult2$Spore_load2+1)
# response<-log10(mtemp.adult2$Dev_pupa)
# 
# #inoculation status + spore load(including many 0) n=325
# m1_mixed<-gam(response ~ s(Temp, k=5) + s(Temp, k=5, by=inoc_status) + s(intensity) + Sex_f + s(Lin_f,Temp, bs="re"), data=mtemp.adult2, method = "REML") #Temp, lineage, sex = sig; inoc_status - no sig interaction with temp

#development to pupation
response<-log10(mtemp.adult$Dev_pupa)

#inoculation status n=332
m1_mixed<-gam(response ~ s(Temp, k=5) + s(Temp, k=5, by=inoc_status) + Sex_f + s(Lin_f,Temp, bs="re"), data=mtemp.adult, method = "REML") #temp, temp-inoc interaction, sex, and lineage are all significant

gam.check(m1_mixed)
summary(m1_mixed)

#spore load n=175
response<-log10(mtemp.adult2$Dev_pupa)
m2_mixed<-gam(response ~ s(Temp, k=4) + s(intensity4) + Sex_f + s(Lin_f,Temp, bs="re"), data=mtemp.adult2, method = "REML") #spore load does not predict development


gam.check(m2_mixed)
summary(m2_mixed)


plot.new()
plot_smooth(m1_mixed, view = "Temp", plot_all = "inoc_status", rm.ranef = FALSE, col=c("purple", "black"), lwd=2, rug = TRUE, ylab=("Development time to pupa"), xlab=("Temperature (°C)"), hide.label = TRUE)
axis(1, at=c(15,35), labels=FALSE, lwd.ticks=0)

plot.new()
plot_smooth(m2_mixed, view = "intensity4",  rm.ranef = FALSE, lwd=2, rug = TRUE, ylab=("Development to pupa"), xlab=("log10 spore load"), hide.label = TRUE)
axis(1, at=c(15,35), labels=FALSE, lwd.ticks=0)
#axis(1, at=seq(18 , 34, by=4), lwd=0, lwd.ticks=1)

```

##Survival
```{r}
#whole dataset
mtemp.surv<-completeFun(mtemp.comp, "Surv_adult")
mtemp.surv$inf_status<-as.numeric(mtemp.surv$Infection_status)
mtemp.surv$logSpore_load2<-log10(mtemp.surv$Spore_load2+1)
response<-mtemp.surv$Surv_adult

m1_mixed<-gam(response ~ s(Temp, k=5) + s(Temp, k=5, by=inoc_status) + s(Lin_f, bs="re"), data=mtemp.surv, method = "REML", family=binomial) #Temp and lineage still important; inoc status*temperature not significant

gam.check(m1_mixed)
summary(m1_mixed)

#spore load n=175 - k is too low for model to fit
# mtemp.surv2<-filter(mtemp.surv, !is.na(Spore_load2)) #look at just monarchs with values for spore load
# response<-as.factor(mtemp.surv2$Surv_pupa)
# 
# m2_mixed<-gam(response ~ s(Temp, k=5) + s(logSpore_load2) + s(Lin_f, Temp, bs="re"), data=mtemp.surv2, method = "REML", family=binomial)
# 
# gam.check(m2_mixed)
# summary(m2_mixed)


data_probsurv <-  
  mtemp.comp %>% 
  drop_na(Surv_adult) %>%   
  group_by(inoc_status) %>% 
  dplyr::summarise(propsurv = sum(as.numeric(Surv_adult), na.rm=T)/n(),
            sepropsurv = sqrt(propsurv*(1-propsurv)/n()),
            n = n())

#########################################
plot.new()
plot_smooth(m1_mixed, view = "Temp", plot_all = "inoc_status", rm.ranef = FALSE, transform = plogis, col=c("purple", "black"), lwd=2, rug = TRUE, ylab=("Survival Probability"), xlab=("Temperature (°C)"), hide.label = TRUE)
axis(1, at=c(15,35), labels=FALSE, lwd.ticks=0)

```

##Wing area
```{r}
mtemp.adult1<-filter(mtemp.adult1, !is.na(Area))
Bleed_stage_f<-as.factor(mtemp.adult1$Bleed_stage.x)
response<-mtemp.adult1$Area

#inoc status
m1_mixed<-gam(response ~ s(Temp, k=5) + s(Temp, k=5, by=inoc_status) + Sex_f + Bleed_stage_f + s(Lin_f, Temp, bs="re"), data=mtemp.adult1, method = "REML") #bleed stage becomes important (none= much larger than larval; pupal = somewhat larger); sex = sig, Temp and lin = sig; inoc_status interaction = not sig

summary(m1_mixed)
gam.check(m1_mixed)

#spore load
mtemp.adult3<-filter(mtemp.adult1, !is.na(Spore_load))
response<-mtemp.adult3$Area

m2_mixed<-gam(response ~ s(Temp, k=4) + s(logSpore_load) + Sex_f + Bleed_stage_f + s(Lin_f, Temp, bs="re"), data=mtemp.adult3, method = "REML") #no correlation between spore load and wing area


gam.check(m2_mixed)
summary(m2_mixed)


plot.new()
plot_smooth(m1_mixed, view = "Temp", plot_all = "inoc_status", rm.ranef = FALSE, col=c("purple", "black"), lwd=2, rug = TRUE, ylab=("Wing area (mm^2)"), xlab=("Temperature (°C)"), hide.label = TRUE)
axis(1, at=c(15,35), labels=FALSE, lwd.ticks=0)

plot.new()
plot_smooth(m2_mixed, view = "logSpore_load", rm.ranef = FALSE, col=c("blue"), lwd=2, rug = TRUE, ylab=("Wing area (mm^2)"), xlab=("log10 spore load"), hide.label = TRUE)
axis(1, at=c(15,35), labels=FALSE, lwd.ticks=0)

################

```

##Adult longevity
```{r}
response<-(mtemp.adult$Adult_longevity)

longev_sum<-
  mtemp.adult %>% 
  group_by(Temp, inoc_status) %>% 
  dplyr::summarise(avg_life = mean(Adult_longevity, na.rm=T))

#inoc status
m1_mixed<-gam(response ~ s(Temp, k=5) + s(Temp, k=5, by=inoc_status)+ Sex_f + s(Lin_f, Temp, bs="re"), data=mtemp.adult, method = "REML", family = 'nb', link='log') #Temp and temp*inoc interaction = sig; inoculated monarchs = consistently lower lifespan, but at 34C both 0 and 1 inoc have low longeveity even though none were infected

overdisp_fun(m1_mixed) #poisson was overdispersed, looks ok with neg-bin
summary(m1_mixed)
gam.check(m1_mixed)

#spore load - we already know this impacts longevity

plot.new()
plot_smooth(m1_mixed, view = "Temp", plot_all = "inoc_status", transform = exp, rm.ranef = FALSE, col=c("purple", "black"), lwd=2, rug = TRUE, ylab=("Adult longevity (days))"), xlab=("Temperature (°C)"), hide.label = TRUE)
axis(1, at=c(15,35), labels=FALSE, lwd.ticks=0)

# plot.new()
# plot_smooth(m2_mixed, view = "logSpore_load", transform = exp, rm.ranef = FALSE, col=c("purple", "black"), lwd=2, rug = TRUE, ylab=("Adult longevity (days))"), xlab=("Temperature (°C)"), hide.label = TRUE)
# axis(1, at=c(15,35), labels=FALSE, lwd.ticks=0)
```

#spore load-longevity relationship
GLM
response = longevity
explanatory = spore load*temperature interaction, sex, OE, lineage
family = negative binomial (MASS package)
```{r}
mtemp.inf$Sex_f<-as.factor(mtemp.inf$Sex)
mtemp.inf1<-filter(mtemp.inf, !is.na(Sex_f) & !is.na(logSpore_load))
mtemp.inf1<-filter(mtemp.inf, ID!=237)#remove spore load of 0 indiv
mtemp.inf1$Temp_f<-as.factor(mtemp.inf1$Temp)
response<-mtemp.inf1$Adult_longevity

m0<-glm.nb(response ~ 1, data = mtemp.inf1)
m1<-glm.nb(response ~ Temp + logSpore_load + Lineage, data = mtemp.inf1)
m2<-glm.nb(response ~ Temp*logSpore_load + Lineage + Sex + OE_strain, data = mtemp.inf1)
m3<-glm.nb(response ~ Temp*logSpore_load, data = mtemp.inf1)
m4<-glm.nb(response ~ logSpore_load, data = mtemp.inf1)

AIC(m0,m1,m2,m3,m4)#used full model (m2 used in paper)

#overdisp_fun(m2)
summary(m2)
tab_model(m2, transform = NULL, show.stat = TRUE, show.se = TRUE, show.ci = FALSE)
anova(m2)
plot(m2)


cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

p1<-ggplot(mtemp.inf1, aes(x = logSpore_load, y = response, group = Temp_f, color = Temp_f, fill=Temp_f)) +
  labs(x=bquote("Spore load "(log[10](spores/monarch))), y = "Adult longevity (days)") +
  geom_smooth(method = 'glm') +
  geom_point(aes(color =Temp_f), alpha = .5)+
  theme_classic()

p1+
  scale_color_brewer(palette = "Dark2")+
  scale_fill_brewer(palette = "Dark2")+
  guides(fill=guide_legend(title = "Lineage"),color=guide_legend(title = "Lineage"))


p2<-ggplot(mtemp.inf1, aes(x = log10(Spore_load), y = response, group = Lin_f, color = Lin_f, fill=Lin_f)) +
  labs(x =bquote("Spore load "(log[10](spores/monarch))), y = "Adult longevity (days)") +
  geom_smooth(method = 'glm', se=FALSE) +
  geom_point(aes(color =Lin_f), alpha = .5)+
  theme_classic()

p2+
  scale_color_brewer(palette = "Dark2")+
  scale_fill_brewer(palette = "Dark2")+
  guides(fill=guide_legend(title = "Lineage"),color=guide_legend(title = "Lineage"))

```



