---
title: "GLM_immuneTradeoffs_July22"
author: "Isabella Ragonese"
date: '2022-07-05'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load libraries, include=FALSE}
library(tidyverse) 
library(ggplot2) 
library(lme4)
library(visreg)
library(lme4)
library(plyr)
library(mgcv)
library(tidymv)
library(voxel)
library(MASS)
library(mgcViz)
library(itsadug)
library(sjPlot)
library(MuMIn)
```

#Bolker et al.
overdispersion function
```{r}
overdisp_fun <- function(model) {
    rdf <- df.residual(model)
    rp <- residuals(model,type="pearson")
    Pearson.chisq <- sum(rp^2)
    prat <- Pearson.chisq/rdf
    pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
    c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}
```


```{r, read in data and create subsets for control monarchs, pupal and larval immune measures, include=FALSE}
mtemp.comp<-read.csv("data/mtemp_comp_Nov1_2021.csv") #master data

mtemp.comp$Spore_load2<-ifelse(mtemp.comp$Infection_status != 0, mtemp.comp$Spore_load, 0) #add column that includes 0 spore load for adults that emerged without parasites
mtemp.comp$inoc_status<-ifelse(mtemp.comp$OE_strain != "control", 1, 0) #add column that includes 0 spore load for adults that emerged without parasites
#mtemp.comp$Temp2<-(mtemp.comp$Temp)^2 #adds column for temp squared to check quadratic term in models
#mtemp.comp$Temp2_s<-scale(mtemp.comp$Temp2) #scales quadratic term to assist with model convergence
#create factor variables for GAM
mtemp.comp$Lin_f<-as.factor(mtemp.comp$Lineage)
mtemp.comp$Sex_f<-as.factor(mtemp.comp$Sex)
mtemp.comp$inoc_f<-as.factor(mtemp.comp$inoc_status)
mtemp.comp$OE_f<-as.factor(mtemp.comp$OE_strain)
mtemp.comp$Wing_deformity_bin<-if_else(mtemp.comp$Wing_deformity==0|mtemp.comp$Wing_deformity==1,0,1) #make binary deformed/not deformed

#comprehensive dataframe with just the infected individuals
mtemp.inf <- subset(mtemp.comp, Infection_status==1)
mtemp.inoc <- subset(mtemp.comp, inoc_status==1)
#comprehensive dataframe with just control individuals
mtemp.control <- subset(mtemp.comp, OE_strain=="control", Infection_status!=1) 

#subset comprehensive dataframe to create dataframes with either larval or pupal bled individuals
pupahemo.df = filter(mtemp.comp, Bleed_stage.x=="Pupa")
#Add log10 Hemo for ID#70 - as if 1 cell found
pupahemo.df[28,30]=1.130334

larvahemo.df = filter(mtemp.comp, Bleed_stage.x=="Larva")

#take out rows with NA for loghemo, or final absorbance (some individuals didn't actually get bled) in larvahemo, pupalhemo, and po datasets
completeFun <- function(data, desiredCols) {
  completeVec <- complete.cases(data[, desiredCols])
  return(data[completeVec, ])
}

larvahemo.df<-completeFun(larvahemo.df, "LogHemo")
larvahemo.df = filter(larvahemo.df, !is.na(Mass))
pupahemo.df<-completeFun(pupahemo.df, "LogHemo")
pupahemo.df = filter(pupahemo.df, !is.na(Mass))

mtemp.po<-completeFun(mtemp.comp, "FinalAbsorbance")
mtemp.adult<-filter(mtemp.comp, Surv_adult==1)
control.adult<-completeFun(mtemp.control, "Dev_adult")


#how many missing values for each column?
control.NA<-colSums(sapply(mtemp.control, FUN=is.na))
larv.na<-colSums(sapply(larvahemo.df, FUN=is.na))

#merozoites #n=46
merozoite_df<-read.csv("data/Merozoites_7_21.csv")
merozoite_df2<-left_join(merozoite_df, mtemp.inf, by="ID")
merozoite_df2$Temp<-as.character(merozoite_df2$Temp)
merozoite_df3<-merozoite_df2[!is.na(merozoite_df2$Temp),]
```

#hemocyte data
```{r}
hemocytes<-read.csv("data/HemocyteCounts_June5.csv")

hemocyteprop <- hemocytes %>% dplyr::select(ID, TotalHemo, G, P, S, O, PropGran, PropSpher)
mtemp.comp2 <- left_join(mtemp.comp, hemocyteprop, by = "ID")

hemocytedf<-filter(mtemp.comp2, !is.na(PropGran.x))
hemocytedf$differential_count<-hemocytedf$P+hemocytedf$O+hemocytedf$G+hemocytedf$S
```

#proportions of immune cells - with bonferroni correction

```{r}
hemocytedf$PlasmProp<-hemocytedf$P/hemocytedf$differential_count
hemocytedf$GranProp<-hemocytedf$G/hemocytedf$differential_count
hemocytedf$OenoProp<-hemocytedf$O/hemocytedf$differential_count
hemocytedf$SpherProp<-hemocytedf$S/hemocytedf$differential_count
hemocytedf$Temp2<-(hemocytedf$Temp)^2

hemocytedf<-filter(hemocytedf, !is.na(Mass))

##Plasmatocytes - alpha = 0.0125
response<-hemocytedf$PlasmProp

m0<-glm(response ~ 1, data=hemocytedf, family=binomial, weights = differential_count)
m1<-glm(response ~ Temp + Temp2 + inoc_status + Mass + Lineage, data=hemocytedf, family=quasibinomial, weights = differential_count)
#Lineage is significant predictor at alpha level

#m1_gam<-gam(response ~ s(Temp,k=5) + inoc_f + s(Mass) + s(Lin_f, Temp, bs="re"), data=hemocytedf, method = "REML", weights= differential_count) 
#Lineage random effect is significant.

summary(m1)
tab_model(m1, transform = NULL, show.stat = TRUE, show.se = TRUE, show.ci = FALSE)
AIC(m0,m1) #AIC is NA for quasibinomial?

##Granulocytes
response<-hemocytedf$GranProp

m0<-glm(response ~ 1, data=hemocytedf, family=binomial, weights = differential_count)
m1<-glm(response ~ Temp + Temp2 + inoc_status + Mass + Lineage, data=hemocytedf, family=quasibinomial, weights = differential_count)
#Just lineage F

# m1_gam<-gam(response ~ s(Temp,k=5) + inoc_f + s(Mass) + s(Lin_f, Temp, bs="re"), data=hemocytedf, method = "REML", weights= differential_count)  #Mass signif
# 
# plot.new()
# plot_smooth(m1_gam, view = "Mass", plot_all = "inoc_f", rm.ranef = FALSE, col=c("purple", "black"), lwd=2, rug = TRUE, ylab=("Proportaion granulocytes"), xlab=("Mass"), hide.label = TRUE)
# axis(1, at=c(0,2), labels=FALSE, lwd.ticks=0)

summary(m1)
tab_model(m1, transform = NULL, show.stat = TRUE, show.se = TRUE, show.ci = FALSE)
AIC(m0,m1)

ggplot(hemocytedf, aes(x=Mass, y=GranProp, fill=inoc_f))+
         labs(x = "Mass (g)", y = "Prop Granulocytes") +
  stat_smooth(aes(group=inoc_f, color=inoc_f), method = 'lm') +
  geom_point(aes(colour = inoc_f), alpha = .5)+
  theme_classic()+
  scale_x_continuous(breaks=c(18,22,26,30,34))

Grantable<-hemocytedf %>% 
  group_by(inoc_f) %>% 
  dplyr::summarise(avgG = mean(GranProp, na.rm=T),
                   seG = sqrt(avgG*(1-avgG)/n()))

ggplot(Grantable, aes(inoc_f, avgG)) +
  geom_point(stat = "identity") +        
  geom_errorbar(aes(ymin=avgG-seG, ymax = avgG+seG), width = .2, 
                position=position_dodge(0.9))

##Spheroids
response<-hemocytedf$SpherProp

m0<-glm(response ~ 1, data=hemocytedf, family=binomial, weights = differential_count)
m1<-glm(response ~ Temp + Temp2 + inoc_status + Mass + Lineage, data=hemocytedf, family=quasibinomial, weights = differential_count)
#only Lineage is significant

#m1_gam<-gam(response ~ s(Temp,k=5) + inoc_f + s(Mass) + s(Lin_f, Temp, bs="re"), data=hemocytedf, method = "REML", weights= differential_count)  #Mass signif

# plot.new()
# plot_smooth(m1_gam, view = "Mass", plot_all = "inoc_f", rm.ranef = FALSE, col=c("purple", "black"), lwd=2, rug = TRUE, ylab=("Proportaion granulocytes"), xlab=("Mass"), hide.label = TRUE)
# axis(1, at=c(0,2), labels=FALSE, lwd.ticks=0)

summary(m1)
tab_model(m1, transform = NULL, show.stat = TRUE, show.se = TRUE, show.ci = FALSE)
AIC(m0,m1)

##Oenocytoids
response<-hemocytedf$OenoProp

m0<-glm(response ~ 1, data=hemocytedf, family=binomial, weights = differential_count)
m1<-glm(response ~ Temp + Temp2 + inoc_status + Mass + Lineage, data=hemocytedf, family=quasibinomial, weights = differential_count)
#no sinif predictors

summary(m1)
tab_model(m1, transform = NULL, show.stat = TRUE, show.se = TRUE, show.ci = FALSE)
AIC(m0,m1)

ggplot(hemocytedf, aes(x=Temp, y=OenoProp))+
         labs(x = "Temperature (°C)", y = "Prop Oenocytoids") +
  stat_smooth(method = 'loess', span=0.9) +
  geom_point(alpha = .5)+
  theme_classic()+
  scale_x_continuous(breaks=c(18,22,26,30,34))

Oenotable<-hemocytedf %>% 
  group_by(Temp) %>% 
  dplyr::summarise(avgO = mean(OenoProp, na.rm=T),
                   seO = sqrt(avgO*(1-avgO)/n()))

ggplot(Oenotable, aes(Temp, avgO)) +
  geom_point(stat = "identity") +        
  geom_errorbar(aes(ymin=avgO-seO, ymax = avgO+seO), width = .2, 
                position=position_dodge(0.9))
```



#Trade-offs between immune capacity and fitness?
Adult longevity, survival prob, size, development time, spore load, infection prob.

#create dataframe for all monarchs that were bled at either larval or pupal stage
```{r}
mtemp.bled<-filter(mtemp.comp, Bleed_stage.x!="None")
mtemp.bled$inoc_f<-as.factor(mtemp.bled$inoc_status)
mtemp.bled$Temp_f<-as.factor(mtemp.bled$Temp)

```


```{r}
larvahemo.df$Temp2<-(larvahemo.df$Temp)^2
pupahemo.df$Temp2<-(pupahemo.df$Temp)^2
larvahemo.df1<-filter(larvahemo.df, (!is.na(Infection_status)&!is.na(FinalAbsorbance)))
pupahemo.df1<-filter(pupahemo.df, (!is.na(Infection_status)&!is.na(FinalAbsorbance)))
response<-pupahemo.df1$Adult_longevity

```


##Adult longevity and immunity
  #accounting for infection status, L and P immunity and interactions with temp are not significant predictors of adult lifespan
```{r}
#response<-log10(larvahemo.df$Adult_longevity)
larvahemo.df1<-filter(larvahemo.df, (!is.na(Infection_status)&!is.na(FinalAbsorbance)))
response<-larvahemo.df1$Adult_longevity
#m1<-lmer(response ~ Temp*LogHemo + inoc_status*LogHemo + (1|Lineage), data = larvahemo.df)

m0<-glm(response ~ 1, data = larvahemo.df1, family = negative.binomial(1))
m1<-glm(response ~ Temp*LogHemo + Temp2 + Infection_status + Lineage + Sex, data = larvahemo.df1, family = negative.binomial(1)) #infection status
m1b<-glm(response ~ Temp*LogHemo + Temp2 + Infection_status, data = larvahemo.df1, family = negative.binomial(1)) #infection status
m1c<-glm(response ~ Infection_status, data = larvahemo.df1, family = negative.binomial(1))


#infection status is the strongest predictor of adult longevity. Model with just infection status performs better than null and models with log hemo or final absorbance

m2<-glm(response ~ Temp*FinalAbsorbance + Temp2 + Infection_status + Sex + Lineage, data = larvahemo.df1, family = negative.binomial(1))
m2b<-glm(response ~ Temp*FinalAbsorbance + Temp2 + Infection_status, data = larvahemo.df1, family = negative.binomial(1))
m2c<-glm(response ~ Infection_status, data = larvahemo.df1, family = negative.binomial(1))

summary(m1)
summary(m2c)

tab_model(m1, transform = NULL, show.stat = TRUE, show.se = TRUE, show.ci = FALSE)
tab_model(m2c, transform = NULL, show.stat = TRUE, show.se = TRUE, show.ci = FALSE)

AICc(m0,m1,m1b,m1c)
AICc(m0,m2,m2b,m2c)

ggplot(larvahemo.df1, aes(x = FinalAbsorbance, y = response, fill=inoc_f, linetype=inoc_f)) +
  labs(x = "larval PO activity (absorbance))", y = "log10 adult longevity") +
  stat_smooth(aes(group=inoc_f, color=inoc_f), method = 'glm') +
  geom_point(aes(color = inoc_f), alpha = .5)+
  theme_classic()

ggplot(larvahemo.df1, aes(x = LogHemo, y = response, fill=inoc_f, linetype=inoc_f)) +
  labs(x = "larval hemocytes (log10(cells/ul))", y = "Adult longevity") +
  stat_smooth(aes(group=inoc_f, color=inoc_f), method = 'glm') +
  geom_point(aes(color = inoc_f), alpha = .5)+
  theme_classic()
####################################################################
#response<-log10(pupahemo.df$Adult_longevity)
#m1<-lmer(response ~ Temp*LogHemo + inoc_status*LogHemo + (1|Lineage), data = pupahemo.df)
pupahemo.df1<-filter(pupahemo.df, (!is.na(Infection_status)&!is.na(FinalAbsorbance)))
response<-pupahemo.df1$Adult_longevity


m0<-glm(response ~ 1, data = pupahemo.df1, family = negative.binomial(1))
m1<-glm(response ~ Temp*LogHemo + Temp2 + Infection_status + Lineage + Sex, data = pupahemo.df1, family = negative.binomial(1)) #infection status
m1b<-glm(response ~ Temp*LogHemo + Temp2 + Infection_status, data = pupahemo.df1, family = negative.binomial(1)) #infection status
m1c<-glm(response ~ Infection_status, data = pupahemo.df1, family = negative.binomial(1)) #

m2<-glm(response ~ Temp*FinalAbsorbance + Temp2 + Infection_status + Sex + Lineage, data = pupahemo.df1, family = negative.binomial(1))
m2b<-glm(response ~ Temp*FinalAbsorbance + Temp2 + Infection_status, data = pupahemo.df1, family = negative.binomial(1))
m2c<-glm(response ~ Infection_status, data = pupahemo.df1, family = negative.binomial(1))
#infection status is the strongest predictor of adult longevity. Model with just infection status performs better than null and models with log hemo or final absorbance

summary(m1)
summary(m2)

tab_model(m1, transform = NULL, show.stat = TRUE, show.se = TRUE, show.ci = FALSE)
tab_model(m2c, transform = NULL, show.stat = TRUE, show.se = TRUE, show.ci = FALSE)

AICc(m0,m1,m1b,m1c)
AICc(m0,m2,m2b,m2c)

ggplot(pupahemo.df1, aes(x = FinalAbsorbance, y = response, fill=inoc_f, linetype=inoc_f)) +
  labs(x = "Pupal PO activity (absorbance))", y = "Adult longevity") +
  stat_smooth(aes(group=inoc_f, color=inoc_f), method = 'lm') +
  geom_point(aes(color = inoc_f), alpha = .5)+
  theme_classic()

ggplot(pupahemo.df1, aes(x = LogHemo, y = response, fill=inoc_f, linetype=inoc_f)) +
  labs(x = "Pupal hemocytes (log10(cells/ul))", y = "Adult longevity") +
  stat_smooth(aes(group=inoc_f, color=inoc_f), method = 'lm') +
  geom_point(aes(color = inoc_f), alpha = .5)+
  theme_classic()

```

##development and immunity
  #Temp and Temp2 overshadow any affects of immunity on development time; interaction not signif
```{r}
response<-log10(larvahemo.df1$Dev_adult)

m0<-lm(response ~ 1, data = larvahemo.df1)
m1<-lm(response ~ Temp*LogHemo + Temp2 + Infection_status + Sex + Lineage, data = larvahemo.df1)
m1b<-lm(response ~ Temp*LogHemo + Temp2, data = larvahemo.df1)
m1c<-lm(response ~ Temp+ Temp2, data = larvahemo.df1)
m2<-lm(response ~ Temp*FinalAbsorbance + Temp2 + Infection_status + Sex + Lineage, data = larvahemo.df1)
m2b<-lm(response ~ Temp*FinalAbsorbance + Temp2, data = larvahemo.df1)
m2c<-lm(response ~ Temp+ Temp2, data = larvahemo.df1)


tab_model(m1c, transform = NULL, show.stat = TRUE, show.se = TRUE, show.ci = FALSE)
tab_model(m2c, transform = NULL, show.stat = TRUE, show.se = TRUE, show.ci = FALSE)

#only temperature and temp2 are significant

summary(m1c)
summary(m2c)

AICc(m0,m1,m1b,m1c)
AICc(m0,m2,m2b,m2c)

ggplot(larvahemo.df1, aes(x = FinalAbsorbance, y = response, fill=inoc_f, linetype=inoc_f)) +
  labs(x = "larval PO activity (absorbance))", y = "log10 development") +
  stat_smooth(aes(group=inoc_f, color=inoc_f), method = 'lm') +
  geom_point(aes(color = inoc_f), alpha = .5)+
  theme_classic()

ggplot(larvahemo.df1, aes(x = LogHemo, y = response, fill=inoc_f, linetype=inoc_f)) +
  labs(x = "larval hemocytes (log10(cells/ul))", y = "log10 development") +
  stat_smooth(aes(group=inoc_f, color=inoc_f), method = 'lm') +
  geom_point(aes(color = inoc_f), alpha = .5)+
  theme_classic()
####################################################################
response<-log10(pupahemo.df1$Dev_adult)
pupahemo.df1$Temp_f<-as.factor(pupahemo.df1$Temp)

m0<-lm(response ~ 1, data = pupahemo.df1)
m1<-lm(response ~ Temp*LogHemo + Temp2 + Infection_status + Sex + Lineage, data = pupahemo.df1)
m1b<-lm(response ~ Temp*LogHemo + Temp2, data = pupahemo.df1)
m1c<-lm(response ~ Temp + Temp2, data = pupahemo.df1)

m2<-lm(response ~ Temp*FinalAbsorbance + Temp2 + Infection_status + Sex + Lineage, data = pupahemo.df1)
m2b<-lm(response ~ Temp*FinalAbsorbance + Temp2, data = pupahemo.df1)
m2c<-lm(response ~ Temp + Temp2, data = pupahemo.df1)

summary(m1b)#best - temp still overwhelmingly important. interaction and hemocyte concentration may explain some more of the variation? but not significant
summary(m2b) #same with PO

AICc(m0,m1,m1b,m1c)
AICc(m0,m2,m2b,m2c)

tab_model(m1, transform = NULL, show.stat = TRUE, show.se = TRUE, show.ci = FALSE)
tab_model(m2, transform = NULL, show.stat = TRUE, show.se = TRUE, show.ci = FALSE)


ggplot(pupahemo.df1, aes(x = FinalAbsorbance, y = response, fill=Temp_f, linetype=Temp_f)) +
  labs(x = "Pupal PO activity (absorbance))", y = "log10 development") +
  stat_smooth(aes(group=Temp_f, color=Temp_f), method = 'lm') +
  geom_point(aes(color = Temp_f), alpha = .5)+
  theme_classic()

ggplot(pupahemo.df1, aes(x = LogHemo, y = response, fill=Temp_f, linetype=Temp_f)) +
  labs(x = "Pupal hemocytes (log10(cells/ul))", y = "log10 development") +
  stat_smooth(aes(group=Temp_f, color=Temp_f), method = 'lm') +
  geom_point(aes(color = Temp_f), alpha = .5)+
  theme_classic()
```

#Size and immunity
  #no affects of immunity or interactions with temp on wing area
```{r}
response<-larvahemo.df1$Area

m0<-lm(response ~ 1, data = larvahemo.df1)
m1<-lm(response ~ Temp*LogHemo + Temp2 + Infection_status + Sex + Lineage, data = larvahemo.df1)
m1b<-lm(response ~ Temp*LogHemo + Temp2 + Infection_status, data = larvahemo.df1)
m1c<-lm(response ~ Temp*LogHemo + Temp2, data = larvahemo.df1)
m2<-lm(response ~ Temp*FinalAbsorbance + Temp2 + Infection_status + Sex + Lineage, data = larvahemo.df1)
m2b<-lm(response ~ Temp*FinalAbsorbance + Temp2 + Infection_status, data = larvahemo.df1)
m2c<-lm(response ~ Temp*FinalAbsorbance + Temp2, data = larvahemo.df1)

AICc(m0,m1,m1b,m1c)
AICc(m0,m2,m2b,m2c)


summary(m1c) #only Temp2 is signif
summary(m2c) #only temp is signif

tab_model(m1c, transform = NULL, show.stat = TRUE, show.se = TRUE, show.ci = FALSE)
tab_model(m2c, transform = NULL, show.stat = TRUE, show.se = TRUE, show.ci = FALSE)

ggplot(larvahemo.df1, aes(x = FinalAbsorbance, y = response, fill=inoc_f, linetype=inoc_f)) +
  labs(x = "larval PO activity (absorbance))", y = "Area") +
  stat_smooth(aes(group=inoc_f, color=inoc_f), method = 'lm') +
  geom_point(aes(color = inoc_f), alpha = .5)+
  theme_classic()

ggplot(larvahemo.df1, aes(x = LogHemo, y = response, fill=inoc_f, linetype=inoc_f)) +
  labs(x = "larval hemocytes (log10(cells/ul))", y = "Area") +
  stat_smooth(aes(group=inoc_f, color=inoc_f), method = 'lm') +
  geom_point(aes(color = inoc_f), alpha = .5)+
  theme_classic()


####################################################################
pupahemo.df2<-filter(pupahemo.df1, !is.na(Area))
response<-pupahemo.df2$Area

Area_temp <-  
  pupahemo.df %>% 
  group_by(Temp) %>% 
  dplyr::summarise(avgarea = mean(Area, na.rm=T),
                   avgFA = mean(FinalAbsorbance, na.rm=T),
            n = n()) #None of the 34C pupae (n=11) with FinalAbsorbance values have area measures

m0<-lm(response ~ 1, data = pupahemo.df2)
m1<-lm(response ~ Temp*LogHemo + Temp2 + Infection_status + Lineage + Sex, data = pupahemo.df2)
#m1<-lmer(response ~ Temp*LogHemo + Temp2 + Infection_status + Sex +(1|Lineage), data = pupahemo.df2) #lineage as random effect
m1b<-lm(response ~ Temp*LogHemo + Temp2 + Infection_status, data = pupahemo.df2)
m1c<-lm(response ~ Temp*LogHemo + Temp2, data = pupahemo.df2)

m2<-lm(response ~ Temp*FinalAbsorbance + Temp2 + Infection_status + Lineage + Sex, data = pupahemo.df2)
m2b<-lm(response ~ Temp*FinalAbsorbance + Temp2 + Infection_status, data = pupahemo.df2)
m2c<-lm(response ~ Temp*FinalAbsorbance + Temp2, data = pupahemo.df2)

AICc(m0,m1,m1b,m1c)
AICc(m0,m2,m2b,m2c)

summary(m1) #lineage and sex are significant predictors, nothing else
summary(m2)

tab_model(m1, transform = NULL, show.stat = TRUE, show.se = TRUE, show.ci = FALSE)

ggplot(pupahemo.df2, aes(x = FinalAbsorbance, y = response, fill=Infection_status)) +
  labs(x = "Pupal PO activity (absorbance))", y = "Area") +
  stat_smooth(aes(group=Infection_status, color=Infection_status), method = 'lm') +
  geom_point(aes(color = Infection_status), alpha = .5)+
  theme_classic()

ggplot(pupahemo.df2, aes(x = FinalAbsorbance, y = response, fill=Temp_f)) +
  labs(x = "Pupal PO activity (absorbance))", y = "Area") +
  stat_smooth(aes(group=Temp_f, color=Temp_f), method = 'lm') +
  geom_point(aes(color = Temp_f), alpha = .5)+
  theme_classic()

ggplot(pupahemo.df2, aes(x = LogHemo, y = response, fill=Temp_f, linetype=Temp_f)) +
  labs(x = "Pupal hemocytes (log10(cells/ul))", y = "Area") +
  stat_smooth(aes(group=Temp_f, color=Temp_f), method = 'lm') +
  geom_point(aes(color = Temp_f), alpha = .5)+
  theme_classic()

ggplot(pupahemo.df2, aes(x = LogHemo, y = response, fill=inoc_f, linetype=inoc_f)) +
  labs(x = "Pupal hemocytes (log10(cells/ul))", y = "Area") +
  stat_smooth(aes(group=inoc_f, color=inoc_f), method = 'lm') +
  geom_point(aes(color = inoc_f), alpha = .5)+
  theme_classic()


```

##try using entire dataset and use bleed stage as term?
```{r}
mtemp.bled<-filter(mtemp.comp, Bleed_stage.x!="None")
mtemp.bled$Temp_f<-as.factor(mtemp.bled$Temp)
response<-mtemp.bled$Area
m1<-lm(response ~ Temp*LogHemo + LogHemo*Bleed_stage.x + Lineage + Sex, data = mtemp.bled)
```

#survival prob and immunity 
  ##
  #No impact of immune measures or their interactions with temp on survival
```{r}
response<-larvahemo.df$Surv_adult
larvahemo.df$Temp_f<-as.factor(larvahemo.df$Temp)
larvahemo.df2<-filter(larvahemo.df, !is.na(FinalAbsorbance))

surv_immun <-  
  larvahemo.df %>% 
  group_by(Temp) %>% 
  dplyr::summarise(propsurv = sum(Surv_adult, na.rm=T)/n(),
                   avgFA = mean(FinalAbsorbance, na.rm=T),
                   avghemo = mean(LogHemo, na.rm=T),
            n = n()) #None of the 34C pupae (n=11) with FinalAbsorbance values have area measures


m0<-glm(response ~ 1, data = larvahemo.df, family = binomial)
m1<-glm(response ~ Temp*LogHemo + Temp2 + inoc_status + Lineage, data = larvahemo.df, family = binomial)
m1b<-glm(response ~ Temp*LogHemo + Temp2 + inoc_status, data = larvahemo.df, family = binomial)
m1c<-glm(response ~ Temp*LogHemo + Temp2, data = larvahemo.df, family = binomial)

tab_model(m1c, transform = NULL, show.stat = TRUE, show.se = TRUE, show.ci = FALSE)
AICc(m0,m1,m1b,m1c)

response<-larvahemo.df2$Surv_adult
m0<-glm(response ~ 1, data = larvahemo.df2, family = binomial)
m2<-glm(response ~ Temp*FinalAbsorbance + Temp2 + inoc_status + Lineage, data = larvahemo.df2, family = binomial)
m2b<-glm(response ~ Temp*FinalAbsorbance + Temp2 + inoc_status, data = larvahemo.df2, family = binomial)
m2c<-glm(response ~ Temp*FinalAbsorbance + Temp2, data = larvahemo.df2, family = binomial)

tab_model(m2c, transform = NULL, show.stat = TRUE, show.se = TRUE, show.ci = FALSE)
AICc(m0,m2,m2b,m2c)

summary(m1c)
summary(m2c)

tab_model(m1c, transform = NULL, show.stat = TRUE, show.se = TRUE, show.ci = FALSE)
tab_model(m2c, transform = NULL, show.stat = TRUE, show.se = TRUE, show.ci = FALSE)

ggplot(larvahemo.df2, aes(x = FinalAbsorbance, y = response, fill=Temp_f, linetype=Temp_f)) +
  labs(x = "larval PO activity (absorbance))", y = "surv prob") +
  stat_smooth(aes(group=Temp_f, color=Temp_f), method = 'lm') +
  geom_point(aes(color = Temp_f), alpha = .5)+
  theme_classic()
ggplot(larvahemo.df2, aes(x = FinalAbsorbance, y = response, fill=inoc_f, linetype=inoc_f)) +
  labs(x = "larval PO activity (absorbance))", y = "surv prob") +
  stat_smooth(aes(group=inoc_f, color=inoc_f), method = 'lm') +
  geom_point(aes(color = inoc_f), alpha = .5)+
  theme_classic()

ggplot(larvahemo.df2, aes(x = LogHemo, y = response, fill=Temp_f, linetype=Temp_f)) +
  labs(x = "larval hemocytes (log10(cells/ul))", y = "surv prob") +
  stat_smooth(aes(group=Temp_f, color=Temp_f), method = 'lm') +
  geom_point(aes(color = Temp_f), alpha = .5)+
  theme_classic()
####################################################################
response<-pupahemo.df$Surv_adult
pupahemo.df$Temp_f<-as.factor(pupahemo.df$Temp)

m0<-glm(response ~ 1, data = pupahemo.df, family=binomial)
m1<-glm(response ~ Temp*LogHemo + Temp2 + inoc_status + Lineage, data = pupahemo.df, family=binomial)
m1b<-glm(response ~ Temp*LogHemo + Temp2 + inoc_status, data = pupahemo.df, family=binomial)
m1c<-glm(response ~ Temp*LogHemo + Temp2, data = pupahemo.df, family=binomial)

AICc(m0,m1,m1b,m1c)

pupahemo.df3<-filter(pupahemo.df, !is.na(FinalAbsorbance))
response<-pupahemo.df3$Surv_adult
m0<-glm(response ~ 1, data = pupahemo.df3, family=binomial)
m2<-glm(response ~ Temp*FinalAbsorbance + Temp2 + inoc_status + Lineage, data = pupahemo.df3, family=binomial)
m2b<-glm(response ~ Temp*FinalAbsorbance + Temp2 + inoc_status, data = pupahemo.df3, family=binomial)
m2c<-glm(response ~ Temp*FinalAbsorbance + Temp2, data = pupahemo.df3, family=binomial)


AICc(m0,m2,m2b,m2c)

summary(m0)#no model is better than the null model for log hemo
summary(m2b) #best model, but no signif predictors (survival may be impacted by temp2 and inoc status)

tab_model(m2b, transform = NULL, show.stat = TRUE, show.se = TRUE, show.ci = FALSE)


ggplot(pupahemo.df3, aes(x = FinalAbsorbance, y = response, fill=Temp_f, linetype=Temp_f)) +
  labs(x = "Pupal PO activity (absorbance))", y = "Surv prob") +
  stat_smooth(aes(group=Temp_f, color=Temp_f), method = 'lm') +
  geom_point(aes(color = Temp_f), alpha = .5)+
  theme_classic()

ggplot(pupahemo.df3, aes(x = FinalAbsorbance, y = response, fill=inoc_f, linetype=inoc_f)) +
  labs(x = "Pupal PO activity (absorbance))", y = "Surv prob") +
  stat_smooth(aes(group=inoc_f, color=inoc_f), method = 'lm') +
  geom_point(aes(color = inoc_f), alpha = .5)+
  theme_classic()

ggplot(pupahemo.df3, aes(x = LogHemo, y = response, fill=Temp_f, linetype=Temp_f)) +
  labs(x = "Pupal hemocytes (log10(cells/ul))", y = "surv prob") +
  stat_smooth(aes(group=Temp_f, color=Temp_f), method = 'lm') +
  geom_point(aes(color = Temp_f), alpha = .5)+
  theme_classic()
```


#immune measures and spore load
  #immunity at P or L stage does not predict infection severity and the relationship does not vary across the 4 temp treatments
```{r}
response<-log10(larvahemo.df$Spore_load+1)
#m1<-lmer(response ~ Temp*LogHemo + inoc_status*LogHemo + (1|Lineage), data = larvahemo.df)

m0<-glm(response ~ 1, data = larvahemo.df)
m1<-glm(response ~ Temp*LogHemo + Temp2 + Sex + Lineage, data = larvahemo.df)
m1b<-glm(response ~ Temp*LogHemo + Temp2, data = larvahemo.df)

AICc(m0,m1,m1b)

response<-log10(larvahemo.df2$Spore_load+1)
m0<-glm(response ~ 1, data = larvahemo.df2)
m2<-glm(response ~ Temp*FinalAbsorbance + Temp2 + Sex + Lineage, data = larvahemo.df2)
m2b<-glm(response ~ Temp*FinalAbsorbance + Temp2, data = larvahemo.df2)

AICc(m0,m2,m2b)

summary(m1) #neither model better than null
summary(m2) #neither model better than null

ggplot(larvahemo.df2, aes(x = FinalAbsorbance, y = response, fill=Temp)) +
  labs(x = "Larval PO activity (absorbance))", y = "log10 spore load") +
  stat_smooth(aes(group=Temp, color=Temp), method = 'glm') +
  geom_point(aes(color = Temp), alpha = .5)+
  theme_classic()

ggplot(larvahemo.df2, aes(x = LogHemo, y = response, fill=Temp)) +
  labs(x = "Larval hemocyte concentration", y = "log10 spore load") +
  stat_smooth(aes(group=Temp, color=Temp), method = 'glm') +
  geom_point(aes(color = Temp), alpha = .5)+
  theme_classic()

###########################################
#pupae
response<-log10(pupahemo.df2$Spore_load+1)
#m1<-lmer(response ~ Temp*LogHemo + inoc_status*LogHemo + (1|Lineage), data = larvahemo.df)

m0<-glm(response ~ 1, data = pupahemo.df2)
m1<-glm(response ~ Temp*LogHemo + Temp2 + Sex + Lineage, data = pupahemo.df2)
m1b<-glm(response ~ Temp*LogHemo + Temp2, data = pupahemo.df2)

AIC(m0,m1,m1b)

m0<-glm(response ~ 1, data = pupahemo.df2)
m2<-glm(response ~ Temp*FinalAbsorbance + Temp2 + Sex + Lineage, data = pupahemo.df2)
m2b<-glm(response ~ Temp*FinalAbsorbance + Temp2, data = pupahemo.df2)

AICc(m0,m2,m2b)

tab_model(m2, transform = NULL, show.stat = TRUE, show.se = TRUE, show.ci = FALSE)

summary(m1)
summary(m2) #sex and lineage predictive of spore load (not PO or Loghemo or interactions with temp)

ggplot(pupahemo.df2, aes(x = FinalAbsorbance, y = response, fill=Temp)) +
  labs(x = "Pupal PO activity (absorbance))", y = "log10 spore load") +
  stat_smooth(aes(group=Temp, color=Temp), method = 'glm') +
  geom_point(aes(color = Temp), alpha = .5)+
  theme_classic()

ggplot(pupahemo.df2, aes(x = LogHemo, y = response, fill=Temp)) +
  labs(x = "Pupal hemocyte concentration", y = "log10 spore load") +
  stat_smooth(aes(group=Temp, color=Temp), method = 'glm') +
  geom_point(aes(color = Temp), alpha = .5)+
  theme_classic()


```

#immunity and prob infection
#
```{r}
larvahemo.df1<-filter(larvahemo.df, !is.na(Sex) & !is.na(FinalAbsorbance))
response<-larvahemo.df1$Infection_status
larvahemo.df1$Temp_f<-as.factor(larvahemo.df1$Temp)
larvahemo.df2<-subset(larvahemo.df, Temp<34)
#response<-larvahemo.df2$Infection_status
#when only looking at 18-30, no relationship. If 18-34, the 34 drives a relationship between absorbance and prob infection.

m0<-glm(response ~ 1, data = larvahemo.df1, family=binomial)
m1<-glm(response ~ Temp*LogHemo + Temp2 + Sex + Lineage, data = larvahemo.df1, family=binomial)
m1b<-glm(response ~ Temp*LogHemo + Temp2, data = larvahemo.df1, family=binomial)

AICc(m0,m1,m1b)

m2<-glm(response ~ Temp*FinalAbsorbance + Temp2 + Sex + Lineage, data = larvahemo.df1, family=binomial) #?why is higher absorbance associated with increased infection prob? does this indicate that the immune system was activated, but not able to clear infection?
m2b<-glm(response ~ Temp*FinalAbsorbance + Temp2, data = larvahemo.df1, family=binomial)

AIC(m0,m2,m2b)
anova(m0,m2, test="Chisq") #m2 is not better than null hypothesis

summary(m1) #not better than null
summary(m2) #not better than null
tab_model(m2, transform = NULL, show.stat = TRUE, show.se = TRUE, show.ci = FALSE)

ggplot(larvahemo.df1, aes(x = FinalAbsorbance, y = response, fill=Temp_f)) +
  labs(x = "Larval PO activity (absorbance))", y = "Infection probability") +
  stat_smooth(aes(group=Temp_f, color=Temp_f), method = 'glm') +
  geom_point(aes(color = Temp_f), alpha = .5)+
  theme_classic()

ggplot(larvahemo.df1, aes(x = FinalAbsorbance, y = response)) +
  labs(x = "Larval PO activity (absorbance))", y = "Infection probability") +
  stat_smooth(method = 'glm') +
  geom_point(alpha = .5)+
  theme_classic()

ggplot(larvahemo.df1, aes(x = LogHemo, y = response, fill=Temp)) +
  labs(x = "Larval hemocyte concentration", y = "Infection probability") +
  stat_smooth(aes(group=Temp, color=Temp), method = 'glm') +
  geom_point(aes(color = Temp), alpha = .5)+
  theme_classic()

###########################################
#pupae
response<-pupahemo.df2$Infection_status
#m1<-lmer(response ~ Temp*LogHemo + inoc_status*LogHemo + (1|Lineage), data = pupahemo.df)

m0<-glm(response ~ 1, data = pupahemo.df2, family=binomial)
m1<-glm(response ~ Temp*LogHemo + Temp2 + Sex + Lineage, data = pupahemo.df2, family=binomial)
m1b<-glm(response ~ Temp*LogHemo + Temp2, data = pupahemo.df2, family=binomial)

AIC(m0,m1,m1b)

m2<-glm(response ~ Temp*FinalAbsorbance + Temp2 + Sex + Lineage, data = pupahemo.df2, family=binomial)
m2b<-glm(response ~ Temp*FinalAbsorbance + Temp2, data = pupahemo.df2, family=binomial)

AIC(m0,m2,m2b)

summary(m1) #null model is better
summary(m2) #null model is better

ggplot(pupahemo.df2, aes(x = FinalAbsorbance, y = response, fill=Temp)) +
  labs(x = "Pupal PO activity (absorbance))", y = "log10 spore load") +
  stat_smooth(aes(group=Temp, color=Temp), method = 'glm') +
  geom_point(aes(color = Temp), alpha = .5)+
  theme_classic()


ggplot(pupahemo.df2, aes(x = LogHemo, y = response, fill=Temp)) +
  labs(x = "Pupal hemocyte concentration", y = "log10 spore load") +
  stat_smooth(aes(group=Temp, color=Temp), method = 'glm') +
  geom_point(aes(color = Temp), alpha = .5)+
  theme_classic()

```

